---
title: "Programa de Prevención y Control de Arbovirosis | 2026 | Chiapas"
format: 
  dashboard:
    nav-buttons: [twitter, github, facebook]
    logo: https://framework-gb.cdn.gob.mx/gobmx/img/logo_blanco.svg
theme: custom.scss
---

```{r setup}
library(sf)
mapview::mapviewOptions(default = TRUE,
                        basemaps.color.shuffle = FALSE)
```

# [**Hotspots de la Transmisión de Dengue**]{style="font-size:8pt"}

::: panel-tabset

### [**Tuxtla Gutierrez**]{style="color:#990000"}

<html>

<head>



```{=html}
<style>
body {
  margin: 0px;
  width: auto; 
  height: auto;
}

.wrapper {
    display: grid;
    grid-gap: 1px;
    grid-template-columns: 225px 225px 225px 410px;
    grid-template-rows: 95px 95px 95px 130px;
    background-color: #772953;
    color: "#772953";
  }

  .box {
    background-color: #772953;
    color: "#772953";
    border-radius: 2px;
    padding: 5px;
    font-size: 100%;

  }

  .a {
    grid-column: 1 / 4;
    grid-row: 1 / 4;
  }
  .b {
    grid-column: 4 ;
    grid-row: 1 / 3;
  }
  .c {
    grid-column: 4 ;
    grid-row: 3 / 5 ;
  }
  .d {
    grid-column: 3;
    grid-row: 4;
  }
  .e {
    grid-column: 2;
    grid-row: 4;
  }
  .f {
    grid-column: 1;
    grid-row: 4;
  }
</style>
```



</head>

<body>

::: wrapper
::: {.box .a}
```{r hotspots_dengue_tuxtla}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "07", 
                                 locality = "Tuxtla Gutierrez")

# Step 2. Extract the hotspots ####
#hotspots <- denmex::den_hotspots[loc, ]
hotspots <- hotspotsdata::hotspots_hex[loc,]
mapview::mapview(hotspots,
                 zcol = "intensity_gi",
                 layer.name = "Intensidad",
                 label = FALSE,
                 color = "white",
                 lwd = 0.5, 
                 col.regions =  rcartocolor::carto_pal(n = max(hotspots$intensity_gi), 
                                                       name = "OrYel"))

```
:::

::: {.box .b}
```{r forest_plot_tuxtla, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_tuxtla,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto))  |> plotly::layout(showlegend = FALSE)

```
:::

::: {.box .d}
```{r gauge_blocks_tuxtla,fig.align = "center", out.height='100%', out.width = '100%', fig.width= 1.5, fig.height=1}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_tuxtla, fig.align = "center", out.height='100%', out.width = '100%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_tuxtla, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::

</body>

</html>


### [**Tapachula**]{style="color:#990000"}

<html>

<head>







</head>

<body>

::: wrapper
::: {.box .a}
```{r hotspots_dengue_tapachula}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "07", 
                                 locality = "Tapachula de Cordoba y Ordoñez")

# Step 2. Extract the hotspots ####
#hotspots <- denmex::den_hotspots[loc, ]
hotspots <- hotspotsdata::hotspots_hex[loc,]
mapview::mapview(hotspots,
                 zcol = "intensity_gi",
                 layer.name = "Intensidad",
                 label = FALSE,
                 color = "white",
                 lwd = 0.5, 
                 col.regions =  rcartocolor::carto_pal(n = max(hotspots$intensity_gi), 
                                                       name = "OrYel"))

```
:::

::: {.box .b}
```{r forest_plot_tapachula, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
 # Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_tapachula,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto))  |> plotly::layout(showlegend = FALSE)

```
:::

::: {.box .d}
```{r gauge_blocks_tapachula,fig.align = "center", out.height='100%', out.width = '100%', fig.width= 1.5, fig.height=1}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_tapachula, fig.align = "center", out.height='100%', out.width = '100%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_tapachula, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::

</body>

</html>

### [**Motozintla**]{style="color:#990000"}

<html>

<head>







</head>

<body>

::: wrapper
::: {.box .a}
```{r hotspots_dengue_motozintla}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "07", 
                                 locality = "Motozintla de Mendoza")

# Step 2. Extract the hotspots ####
hotspots <- denmex::den_hotspots[loc, ]
hotspots <- hotspotsdata::hotspots_hex[loc,]
#hotspots <- hotspotsdata::hotspots_hex[loc,]
mapview::mapview(hotspots,
                 zcol = "intensity_gi",
                 layer.name = "Intensidad",
                 label = FALSE,
                 color = "white",
                 lwd = 0.5, 
                 col.regions =  rcartocolor::carto_pal(n = max(hotspots$intensity_gi), 
                                                       name = "OrYel"))

```
:::

::: {.box .b}
```{r forest_plot_motozintla, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}

# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))

```
:::

::: {.box .c}
```{r power_law_plot_motozintla,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto))  |> plotly::layout(showlegend = FALSE)

```
:::

::: {.box .d}
```{r gauge_blocks_motozintla,fig.align = "center", out.height='100%', out.width = '100%', fig.width= 1.5, fig.height=1}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_motozintla, fig.align = "center", out.height='100%', out.width = '100%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_motozintla, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::

</body>

</html>

### [**Palenque**]{style="color:#990000"}

<html>

<head>






</head>

<body>

::: wrapper
::: {.box .a}
```{r hotspots_dengue_palenque}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "07", 
                                 locality = "Palenque")

# Step 2. Extract the hotspots ####
#hotspots <- denmex::den_hotspots[loc, ]
hotspots <- hotspotsdata::hotspots_hex[loc,]
mapview::mapview(hotspots,
                 zcol = "intensity_gi",
                 layer.name = "Intensidad",
                 label = FALSE,
                 color = "white",
                 lwd = 0.5, 
                 col.regions =  rcartocolor::carto_pal(n = max(hotspots$intensity_gi), 
                                                       name = "OrYel"))

```
:::

::: {.box .b}
```{r forest_plot_palenque, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
 # Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_palenque,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto))  |> plotly::layout(showlegend = FALSE)

```
:::

::: {.box .d}
```{r gauge_blocks_palenque,fig.align = "center", out.height='100%', out.width = '100%', fig.width= 1.5, fig.height=1}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_palenque, fig.align = "center", out.height='100%', out.width = '100%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_palenque, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::

</body>

</html>


### [**Tonalá**]{style="color:#990000"}

<html>

<head>






</head>

<body>

::: wrapper
::: {.box .a}
```{r hotspots_dengue_tonala}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "07", 
                                 locality = "Tonalá")

# Step 2. Extract the hotspots ####
#hotspots <- denmex::den_hotspots[loc, ]
hotspots <- hotspotsdata::hotspots_hex[loc,]
mapview::mapview(hotspots,
                 zcol = "intensity_gi",
                 layer.name = "Intensidad",
                 label = FALSE,
                 color = "white",
                 lwd = 0.5, 
                 col.regions =  rcartocolor::carto_pal(n = max(hotspots$intensity_gi), 
                                                       name = "OrYel"))

```
:::

::: {.box .b}
```{r forest_plot_tonala, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_tonala,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto))  |> plotly::layout(showlegend = FALSE)

```
:::

::: {.box .d}
```{r gauge_blocks_tonala,fig.align = "center", out.height='100%', out.width = '100%', fig.width= 1.5, fig.height=1}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_tonala, fig.align = "center", out.height='100%', out.width = '100%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_tonala, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::

</body>

</html>

### [**Arriaga**]{style="color:#990000"}

<html>

<head>







</head>

<body>

::: wrapper
::: {.box .a}
```{r hotspots_dengue_arriaga}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "07", 
                                 locality = "Arriaga")

# Step 2. Extract the hotspots ####
#hotspots <- denmex::den_hotspots[loc, ]
hotspots <- hotspotsdata::hotspots_hex[loc,]
mapview::mapview(hotspots,
                 zcol = "intensity_gi",
                 layer.name = "Intensidad",
                 label = FALSE,
                 color = "white",
                 lwd = 0.5, 
                 col.regions =  rcartocolor::carto_pal(n = max(hotspots$intensity_gi), 
                                                       name = "OrYel"))

```
:::

::: {.box .b}
```{r forest_plot_arriaga, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_arriagaa,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto))  |> plotly::layout(showlegend = FALSE)

```
:::

::: {.box .d}
```{r gauge_blocks_arriaga,fig.align = "center", out.height='100%', out.width = '100%', fig.width= 1.5, fig.height=1}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_arriaga, fig.align = "center", out.height='100%', out.width = '100%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_arriaga, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::

</body>

</html>


### [**Huixtla**]{style="color:#990000"}

<html>

<head>






</head>

<body>

::: wrapper
::: {.box .a}
```{r hotspots_dengue_huixtla}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "07", 
                                 locality = "Huixtla")

# Step 2. Extract the hotspots ####
#hotspots <- denmex::den_hotspots[loc, ]
hotspots <- hotspotsdata::hotspots_hex[loc,]
mapview::mapview(hotspots,
                 zcol = "intensity_gi",
                 layer.name = "Intensidad",
                 label = FALSE,
                 color = "white",
                 lwd = 0.5, 
                 col.regions =  rcartocolor::carto_pal(n = max(hotspots$intensity_gi), 
                                                       name = "OrYel"))

```
:::

::: {.box .b}
```{r forest_plot_huixtla, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
 
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_huixtla,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto))  |> plotly::layout(showlegend = FALSE)

```
:::

::: {.box .d}
```{r gauge_blocks_huixtla,fig.align = "center", out.height='100%', out.width = '100%', fig.width= 1.5, fig.height=1}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_huixtla, fig.align = "center", out.height='100%', out.width = '100%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_huixtla, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::

</body>

</html>

### [**Ciudad Hidalgo**]{style="color:#990000"}

<html>

<head>







</head>

<body>

::: wrapper
::: {.box .a}
```{r hotspots_dengue_cd_hidalgo}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "07", 
                                 locality = "Ciudad Hidalgo")

# Step 2. Extract the hotspots ####
#hotspots <- denmex::den_hotspots[loc, ]
hotspots <- hotspotsdata::hotspots_hex[loc,]
mapview::mapview(hotspots,
                 zcol = "intensity_gi",
                 layer.name = "Intensidad",
                 label = FALSE,
                 color = "white",
                 lwd = 0.5, 
                 col.regions =  rcartocolor::carto_pal(n = max(hotspots$intensity_gi), 
                                                       name = "OrYel"))

```
:::

::: {.box .b}
```{r forest_plot_cd_hidalgo, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
# Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_cd_hidalgo,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto))  |> plotly::layout(showlegend = FALSE)

```
:::

::: {.box .d}
```{r gauge_blocks_cd_hidalgo,fig.align = "center", out.height='100%', out.width = '100%', fig.width= 1.5, fig.height=1}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_cd_hidalgo, fig.align = "center", out.height='100%', out.width = '100%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_cd_hidalgo, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::

</body>

</html>

### [**Frontera Comalapa**]{style="color:#990000"}

<html>

<head>





</head>

<body>

::: wrapper
::: {.box .a}
```{r hotspots_dengue_comalapa}
# Step 1. extract the locality ####
loc <- rgeomex::extract_locality(cve_edo = "07", 
                                 locality = "Frontera Comalapa")

# Step 2. Extract the hotspots ####
#hotspots <- denmex::den_hotspots[loc, ]
hotspots <- hotspotsdata::hotspots_hex[loc,]
mapview::mapview(hotspots,
                 zcol = "intensity_gi",
                 layer.name = "Intensidad",
                 label = FALSE,
                 color = "white",
                 lwd = 0.5, 
                 col.regions =  rcartocolor::carto_pal(n = max(hotspots$intensity_gi), 
                                                       name = "OrYel"))

```
:::

::: {.box .b}
```{r forest_plot_comalapa, dpi = 300, fig.height= 5,fig.width = 7,fig.align = "center", out.height='100%', out.width = '100%',fig.show = "hold"}
 # Logistic Regression
hotspots <- hotspots |>
    dplyr::mutate(DENV_2008_2010 = DENV_2008 + DENV_2009 + DENV_2010) |>
    dplyr::mutate(DENV_2011_2014 = DENV_2011 + DENV_2012 + DENV_2013 + DENV_2014) |>
    dplyr::mutate(DENV_2015_2017 = DENV_2015 +  DENV_2016 + DENV_2017) |>
    dplyr::mutate(DENV_2018_2019 = DENV_2018 +  DENV_2019) |>
    dplyr::mutate(DENV_2020_2022 = DENV_2020 +  DENV_2021 + DENV_2022) |>
    dplyr::mutate(DENV_2023_2024 = DENV_2023 +  DENV_2024) 


# Logistic Regression
glm.fit <- glm(hotspots_gi~DENV_2008_2010 + DENV_2011_2014 +
                   DENV_2015_2017 + DENV_2018_2019 +
                   DENV_2020_2022 + DENV_2023_2024,
               data = hotspots, 
               family = binomial)

result <- parameters::model_parameters(glm.fit, 
                                       exponentiate = FALSE)

plotly::ggplotly(plot(result) +
       ggplot2::theme(axis.text.y = ggplot2::element_text(size = 8),
                       axis.text.x = ggplot2::element_text(size = 8),
                       axis.title = ggplot2::element_text(size = 8),
                      plot.margin = ggplot2::unit(c(0,0,0,0), "cm")) +
        ggplot2::scale_y_discrete(labels = c("2008-2010",
                                             "2011-2014", 
                                             "2015-2017",
                                             "2018-2019",
                                             "2020-2022",
                                             "2023-2024")))
```
:::

::: {.box .c}
```{r power_law_plot_comalapa,dpi = 300, warning=FALSE, fig.align = "center", out.width = '100%',out.height='100%',fig.show = "hold"}
# Step 7. make the function por calculate the cumulative ####
funcumsum <- function(x){
    x$cumulative_n <- cumsum(x$n_cases)
    x$perc_cumsum_n <- round((cumsum(x$n_cases)/sum(x$n_cases))*100, digits = 1)
    #x$cumulative_ha <- cumsum(x$ha)
    #x$perc_cumsum_ha <- round((cumsum(x$ha)/sum(x$ha))*100, digits = 1)
    x$id <- 1
    x$id_cumsum <- cumsum(x$id)
    x$id_perc_cum <- round((cumsum(x$id)/nrow(x))*100, digits = 1)
    x
}
# step 8.
data_pareto <- hotspots   |>
    # sf::st_drop_geometry()  |>
    # dplyr::mutate(n_cases = rowSums(dplyr::select(., dplyr::starts_with("DENV"))),
    #              loc = "locality")
    #dplyr::mutate(n_cases = rowSums(dplyr::across(DENV_2008:DENV_2021)),
    #              loc = "locality")  |>
    dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    
    dplyr::select(loc, n_cases, hotspots_gi)  |>
    dplyr::arrange(loc, desc(hotspots_gi), desc(n_cases))  |>
    dplyr::group_by(loc)  |>
    tidyr::nest()  |>
    dplyr::mutate(pareto_cases = purrr::map(data,
                                            funcumsum))  |>
    dplyr::select(-data)  |>
    tidyr::unnest(cols = c(pareto_cases))

# step 4. visualization of pareto rules 
plotly::ggplotly(denhotspots::power_law_plot(x = data_pareto))  |> plotly::layout(showlegend = FALSE)

```
:::

::: {.box .d}
```{r gauge_blocks_comalapa,fig.align = "center", out.height='100%', out.width = '100%', fig.width= 1.5, fig.height=1}
y_hot <- hotspots |> dplyr::filter(hotspots_gi == 1)
centroid <- rgeomex::blocks_ine20_mx_centroid |>
  sf::st_as_sf(coords = c("x", "y"), crs = 4326)
b_loc <- centroid[hotspots, ]
b_hot <- b_loc[y_hot, ]
flexdashboard::gauge(round((nrow(b_hot)/nrow(b_loc))*100, digits = 1),
                     min = 0, max = 100, 
                     symbol = '%', 
                     label = "Manzanas",
                     flexdashboard::gaugeSectors(success = c(0, 30), 
                                                 warning = c(31, 50), 
                                                 danger = c(51, 100),
                                                 colors = c("#11BB33", #"#63C1A0", 
                                                            "#FC8427",
                                                            "#E01A59")))
```
:::

::: {.box .e}
```{r gauge_AGEB_comalapa, fig.align = "center", out.height='100%', out.width = '100%'}

y <- hotspots  |>
    sf::st_drop_geometry() |>
     dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)
  
flexdashboard::gauge(round((nrow(y_hot)/nrow(y))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "AGEBs",
      flexdashboard::gaugeSectors(success = c(0, 30), 
                                  warning = c(31, 50), 
                                  danger = c(51, 100),
                                  colors = c("#11BB33",
                                             "#FC8427",
                                             "#E01A59")))
```
:::

::: {.box .f}
```{r gauge_casos_comalapa, out.height='90%', out.width = '90%', fig.align = "center"}

y <- hotspots  |>
    sf::st_drop_geometry() |>
      dplyr::mutate(n_cases = rowSums(dplyr::across(dplyr::starts_with("DENV"))),
                  loc = "locality") |>
    dplyr::select(loc, n_cases, hotspots_gi) 
y_hot <- y |>dplyr::filter(hotspots_gi == 1)

flexdashboard::gauge(round((sum(y_hot$n_cases)/sum(y$n_cases))*100, digits = 1),
      min = 0, max = 100, 
      symbol = '%', 
      label = "Casos de Dengue",
      flexdashboard::gaugeSectors(success = c(11, 100), 
                                  warning = c(6, 10), 
                                  danger = c(0, 5),
                                  colors = c("#11BB33", 
                                             "#FC8427",
                                             "#E01A59")))
```
:::
:::

</body>

</html>




:::




# [**Hotspots del vector del Dengue**]{style="font-size:8pt"}

::: panel-tabset

### [**Tapachula**]{style="color:#990000"}

```{r vector_hotspots_tapachula, dpi = 300, warning=FALSE, message=FALSE}
deneggs::map_eggs_hotspots(betas = denmex::eggs_betas_07_chiapas,
                           locality = "Tapachula de Cordoba Y Ordoñez",
                           cve_edo = "07",
                          static_map = TRUE,
                           palette = rcartocolor::carto_pal,
                           name = "SunsetDark")
```

### [**Motozintla**]{style="color:#990000"}

```{r vector_hotspots_motozintla, dpi = 300, warning=FALSE, message=FALSE}
deneggs::map_eggs_hotspots(betas = denmex::eggs_betas_07_chiapas,
                           locality = "Motozintla de Mendoza",
                           cve_edo = "07",
                           palette = rcartocolor::carto_pal,
                           static_map = TRUE,
                           name = "SunsetDark")
```


:::


# [**Escenarios Epidemiológicos**]{style="font-size:8pt"}


::: panel-tabset

### [**Tapachula**]{style="color:#990000"}

```{r risk_ageb_tapachula}
# Step 1. calculate the risk ####
risk <- denhotspots::risk_ageb(betas = denmex::eggs_betas_07_chiapas,
                               hotspots = denmex::den_hotspots,
                               intensity_perc = 25,
                               locality = "Tapachula de Cordoba Y 0rdoñez",
                               cve_edo = "07")

# Step 2. map ####
denhotspots::map_risk(risk = risk,
                      staticmap = FALSE)

```

### [**Motozintla**]{style="color:#990000"}

```{r risk_ageb_motozintla}
# Step 1. calculate the risk ####
risk <- denhotspots::risk_ageb(betas = denmex::eggs_betas_07_chiapas,
                               hotspots = denmex::den_hotspots,
                               intensity_perc = 25,
                               locality = "Montozintla de Mendoza",
                               cve_edo = "07")

# Step 2. map ####
denhotspots::map_risk(risk = risk,
                      staticmap = FALSE)

```
:::
